---
title: "Grades of Students by School"
output: html_notebook
---


```{r}
grades <- read.csv('data.csv')
```
```{r}
cols <- c("school", "section", "very_ahead", "middling", "behind", "more_behind", "very_behind", "completed")
colnames(grades) <- cols
```

```{r}
attach(grades)
```

```{r}
summary(grades)
```

```{r}
sections_by_school <- aggregate(section, by = list(school), FUN = max)
sections_by_school
```
If the semester is nearly 3/4 of the way over then school C, D, and E must be on different schedules because they've only completed 1 section while A and B have completed at least 10 more.

```{r}
current_course <- max(grades[which(school == "B"), ]$section) / 3 * 4
30 - current_course
```

If one of the schools which has the most sections completed, school B, only has 12 sections done. And the semester is already 3/4 of the way through. They will have to do 14 more sections (2 more than they've completed to date)  in 1/3 of the time -- 1/4 of the total time of the semester. 

This seems odd. Either the final sections of this math course are increasingly easier than the previous 3/4, or they will have to rush the last 1/4 of the semester, or they may not finish. Unless the first 3/4 of the semester are filled with days off for holidays and the previous pace was comparatively slow. Which still would not explain why C, D, and E only have one to three sections done.

(School B was chosen because 12 cleanly makes 3/4 of 16.)


```{r}
maxA <- max(grades[which(school == "A"), ]$section)
maxB <- max(grades[which(school == "B"), ]$section)
grades[which(school == "A" & section == maxA), "completed"]
```

Five kids have finished from school A. 

```{r}
grades[which(school == "B" & section == maxB), "completed"]
```

 19 kids have finished from school B.
 
 
```{r}
max(very_ahead)
```
 
 No kids in any school are 'very ahead'. 
 
```{r}
# Some quick histograms
getHist <- function(vectorX, bin, xlabel = "x", ylabel = "y") {
  qplot(vectorX,
        geom="histogram",
        binwidth = bin,
        fill=I("steelblue"),
        col=I("darkorange"),
        xlab = xlabel,
        ylab = ylabel
        )
}
```

```{r}
getHist(middling, 5, "middling", "frequency")
getHist(very_behind, 1, "very behind", "frequency")
getHist(more_behind, 1, "more behind", "frequency")
```

Not much can be derived from these, there aren't strong patterns here. One stand-out is that one section of one school reported over 20 kids (almost 25) were 'very behind.'
 

```{r}
ggplot(grades, aes(x = school, y = completed)) + geom_boxplot()
```


This is a strange result. If we remember from earlier, School E only has one section finished and they have over 20 kids who have completed the entire course. 


```{r}
ggplot(grades, aes(x = school, y = log(completed))) + geom_boxplot()
```
Changing the completed to log evens out the schools that have finished fewer sections, but shows school E is still an outlier.

```{r}
behind_plot <- ggplot(grades, aes(x = section, y = value, color = variable)) +
  geom_point(aes(y = very_behind, col = "very behind")) +
  geom_point(aes(y = more_behind, col = "more behind")) +
  geom_point(aes(y = behind, col = "behind")) +
  facet_grid(. ~ school)
behind_plot
```

This shows a gradual downtrend among students who are behind each week as the sections progress.

```{r}

doing_well_plot <- ggplot(grades, aes(x = section, y = value, color = variable)) +
  geom_point(aes(y = middling, col = "middling")) +
  geom_point(aes(y = completed, col = "completed")) +
  facet_grid(. ~ school)
doing_well_plot
```

Here there is random noise, but a noticeable uptrend among middling students each section as school B.
